---
title: "AI-READI Exploratory Analysis v2"
author: "Alireza Delfarah"
date: "`r Sys.Date()`"
engine: knitr

format: 
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show R Code"
    theme: flatly
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
# Load required libraries
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(DT)
library(plotly)
library(knitr)
library(kableExtra)
library(scales)
library(tidyr)
library(lubridate)

# Set global options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load column information
col_info <- read_csv("col_df.csv")

# Set minimum days threshold
MIN_DAYS_THRESHOLD <- 5
```

```{r read data}
data_steps <- readRDS("aireadi_steps.rda")
data_hr <- readRDS("aireadi_hr.rda")
data_sleep <- readRDS("aireadi_sleep.rda")
data_stress <- readRDS("aireadi_stress.rda")
data_kcal <- readRDS("aireadi_kcal.rda")
data_rr <- readRDS("aireadi_respiratory.rda")
data_oxygen <- readRDS("aireadi_oxygen.rda")
data_cgm <- readRDS("aireadi_cgm.rda")
data_ecg <- readRDS("aireadi_ecg.rda")
data_clinical <- readRDS("aireadi_clinical.rda")
```

## Overview

This document provides a comprehensive analysis of health data coverage across multiple modalities. The analysis examines:

1. **Number of subjects** present in each data modality (before and after filtering)
2. **Days of data available** per subject in each modality
3. **Subject filtering** based on minimum data requirement (≥`r MIN_DAYS_THRESHOLD` days)
4. **Data missingness patterns** across modalities
5. **Date ranges and temporal coverage** for each participant
6. **Actual participant IDs** instead of generic subject labels

```{r data-analysis-functions}
# Function to extract participant ID from dataframe
extract_participant_id <- function(df, modality_cols) {
  participant_col <- modality_cols[grepl("participant_id", modality_cols, ignore.case = TRUE)][1]
  
  if (!is.null(participant_col) && participant_col %in% names(df) && nrow(df) > 0) {
    # Get the first non-NA participant_id
    participant_ids <- unique(df[[participant_col]][!is.na(df[[participant_col]])])
    if (length(participant_ids) > 0) {
      return(as.character(participant_ids[1]))
    }
  }
  return(NA)
}

# Function to extract date ranges from dataframe
extract_date_range <- function(df, modality_cols) {
  # Find timestamp/date columns
  date_cols <- modality_cols[grepl("date|timestamp|start_time|end_time", modality_cols, ignore.case = TRUE)]
  
  date_info <- list(
    start_date = NA,
    end_date = NA,
    date_column_used = NA,
    unique_dates = 0,
    date_gaps = 0,
    coverage_percentage = NA
  )
  
  if (is.null(df) || nrow(df) == 0 || length(date_cols) == 0) {
    return(date_info)
  }
  
  # Try different date column approaches
  for (date_col in date_cols) {
    if (date_col %in% names(df)) {
      tryCatch({
        if (grepl("date", date_col, ignore.case = TRUE)) {
          # If it's a date column
          dates <- as.Date(df[[date_col]])
        } else {
          # If it's a timestamp, extract date
          timestamps <- as.POSIXct(df[[date_col]])
          dates <- as.Date(timestamps)
        }
        
        # Remove NA dates
        valid_dates <- dates[!is.na(dates)]
        
        if (length(valid_dates) > 0) {
          unique_dates <- unique(valid_dates)
          date_info$start_date <- min(unique_dates)
          date_info$end_date <- max(unique_dates)
          date_info$date_column_used <- date_col
          date_info$unique_dates <- length(unique_dates)
          
          # Calculate date gaps
          if (length(unique_dates) > 1) {
            date_range <- seq(from = min(unique_dates), to = max(unique_dates), by = "day")
            missing_dates <- length(date_range) - length(unique_dates)
            date_info$date_gaps <- missing_dates
            date_info$coverage_percentage <- (length(unique_dates) / length(date_range)) * 100
          } else {
            date_info$coverage_percentage <- 100
          }
          
          break  # Use first successful date column
        }
      }, error = function(e) {
        # Try next column
      })
    }
  }
  
  return(date_info)
}

# Function to calculate missingness for a dataframe
calculate_missingness <- function(df) {
  if (is.null(df) || nrow(df) == 0) {
    return(list(
      total_values = 0,
      missing_values = 0,
      missing_percentage = 100,
      columns_with_missing = character(0),
      column_missingness = data.frame()
    ))
  }
  
  total_values <- nrow(df) * ncol(df)
  missing_values <- sum(is.na(df))
  missing_percentage <- (missing_values / total_values) * 100
  
  # Column-wise missingness
  col_missing <- df %>%
    summarise_all(~sum(is.na(.))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "missing_count") %>%
    mutate(
      total_rows = nrow(df),
      missing_percentage = (missing_count / total_rows) * 100
    ) %>%
    arrange(desc(missing_percentage))
  
  columns_with_missing <- col_missing$column[col_missing$missing_count > 0]
  
  return(list(
    total_values = total_values,
    missing_values = missing_values,
    missing_percentage = missing_percentage,
    columns_with_missing = columns_with_missing,
    column_missingness = col_missing
  ))
}

# Function to analyze data coverage with participant IDs, missingness, and date ranges
analyze_data_coverage <- function(data_list, modality_name, col_info) {
  
  # Get column names for this modality
  modality_cols <- col_info %>% 
    filter(modality == paste0("col_", modality_name)) %>%
    select(-modality) %>%
    unlist() %>%
    na.omit() %>%
    as.character()
  
  # Number of subjects
  n_subjects <- length(data_list)
  
  # Analyze each subject's data
  subject_results <- list()
  modality_missingness <- list()
  
  for (i in 1:length(data_list)) {
    df <- data_list[[i]]
    
    if (!is.null(df) && nrow(df) > 0) {
      
      # Extract participant ID
      participant_id <- extract_participant_id(df, modality_cols)
      if (is.na(participant_id)) {
        participant_id <- paste0("Unknown_", i)
      }
      
      # Calculate missingness
      missingness_info <- calculate_missingness(df)
      modality_missingness[[participant_id]] <- missingness_info
      
      # Extract date range information
      date_info <- extract_date_range(df, modality_cols)
      
      subject_results[[i]] <- list(
        participant_id = participant_id,
        subject_index = i,
        n_records = nrow(df),
        n_days = date_info$unique_dates,
        columns = names(df),
        missingness = missingness_info,
        date_range = date_info
      )
    } else {
      # Handle empty/null dataframes
      participant_id <- paste0("Empty_", i)
      missingness_info <- calculate_missingness(df)
      modality_missingness[[participant_id]] <- missingness_info
      
      date_info <- list(
        start_date = NA,
        end_date = NA,
        date_column_used = NA,
        unique_dates = 0,
        date_gaps = 0,
        coverage_percentage = NA
      )
      
      subject_results[[i]] <- list(
        participant_id = participant_id,
        subject_index = i,
        n_records = 0,
        n_days = 0,
        columns = character(0),
        missingness = missingness_info,
        date_range = date_info
      )
    }
  }
  
  return(list(
    modality = modality_name,
    n_subjects = n_subjects,
    subject_results = subject_results,
    available_columns = modality_cols,
    modality_missingness = modality_missingness
  ))
}

# Function to filter subjects and track removals with participant IDs
filter_subjects_by_days <- function(results, min_days = MIN_DAYS_THRESHOLD) {
  
  filtered_results <- results
  removal_tracking <- list()
  
  for (modality in names(results)) {
    result <- results[[modality]]
    
    # Track original subjects with participant IDs
    original_participants <- sapply(result$subject_results, function(x) x$participant_id)
    original_days <- sapply(result$subject_results, function(x) 
      ifelse(is.na(x$n_days) || x$n_days == 0, 0, x$n_days))
    
    # Filter subjects with sufficient days
    valid_subjects <- original_days >= min_days
    filtered_subject_results <- result$subject_results[valid_subjects]
    
    # Track removals
    removed_participants <- original_participants[!valid_subjects]
    removed_days <- original_days[!valid_subjects]
    
    removal_tracking[[modality]] <- list(
      original_count = length(original_participants),
      filtered_count = length(filtered_subject_results),
      removed_count = length(removed_participants),
      removed_participants = removed_participants,
      removed_days = removed_days,
      removal_rate = length(removed_participants) / length(original_participants)
    )
    
    # Update filtered results
    filtered_results[[modality]]$subject_results <- filtered_subject_results
    filtered_results[[modality]]$n_subjects <- length(filtered_subject_results)
  }
  
  return(list(
    filtered_results = filtered_results,
    removal_tracking = removal_tracking
  ))
}

# Function to create date range summary
create_date_range_summary <- function(results) {
  
  date_range_summary <- data.frame(
    Modality = character(),
    Participant_ID = character(),
    Start_Date = as.Date(character()),
    End_Date = as.Date(character()),
    N_Days = numeric(),
    Date_Gaps = numeric(),
    Coverage_Percentage = numeric(),
    Date_Column_Used = character(),
    Duration_Days = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (modality in names(results$filtered_results)) {
    result <- results$filtered_results[[modality]]
    
    for (subject_result in result$subject_results) {
      participant_id <- subject_result$participant_id
      date_info <- subject_result$date_range
      
      # Calculate duration
      duration_days <- NA
      if (!is.na(date_info$start_date) && !is.na(date_info$end_date)) {
        duration_days <- as.numeric(date_info$end_date - date_info$start_date) + 1
      }
      
      date_range_summary <- rbind(date_range_summary, data.frame(
        Modality = toupper(modality),
        Participant_ID = participant_id,
        Start_Date = as.Date(date_info$start_date),
        End_Date = as.Date(date_info$end_date),
        N_Days = date_info$unique_dates,
        Date_Gaps = date_info$date_gaps,
        Coverage_Percentage = round(date_info$coverage_percentage, 2),
        Date_Column_Used = ifelse(is.na(date_info$date_column_used), "None", date_info$date_column_used),
        Duration_Days = duration_days
      ))
    }
  }
  
  return(date_range_summary)
}

# Function to create temporal overlap analysis
create_temporal_overlap_analysis <- function(date_range_summary) {
  
  # Get overall date range across all modalities
  all_start_dates <- date_range_summary$Start_Date[!is.na(date_range_summary$Start_Date)]
  all_end_dates <- date_range_summary$End_Date[!is.na(date_range_summary$End_Date)]
  
  if (length(all_start_dates) == 0 || length(all_end_dates) == 0) {
    return(NULL)
  }
  
  overall_start <- min(all_start_dates)
  overall_end <- max(all_end_dates)
  
  # Create overlap matrix by participant
  participants <- unique(date_range_summary$Participant_ID)
  modalities <- unique(date_range_summary$Modality)
  
  overlap_matrix <- expand.grid(
    Participant_ID = participants,
    Modality = modalities,
    stringsAsFactors = FALSE
  ) %>%
    left_join(date_range_summary, by = c("Participant_ID", "Modality")) %>%
    mutate(
      Has_Data = !is.na(Start_Date),
      Days_From_Overall_Start = ifelse(Has_Data, as.numeric(Start_Date - overall_start), NA),
      Days_To_Overall_End = ifelse(Has_Data, as.numeric(overall_end - End_Date), NA)
    )
  
  return(list(
    overlap_matrix = overlap_matrix,
    overall_start = overall_start,
    overall_end = overall_end,
    total_study_duration = as.numeric(overall_end - overall_start) + 1
  ))
}

# Main analysis function
run_coverage_analysis <- function() {
  
  # Define the data objects (assuming they are already loaded in environment)
  data_objects <- list(
    cgm = if(exists("data_cgm")) data_cgm else NULL,
    hr = if(exists("data_hr")) data_hr else NULL,
    kcal = if(exists("data_kcal")) data_kcal else NULL,
    oxygen = if(exists("data_oxygen")) data_oxygen else NULL,
    rr = if(exists("data_rr")) data_rr else NULL,
    sleep = if(exists("data_sleep")) data_sleep else NULL,
    steps = if(exists("data_steps")) data_steps else NULL,
    stress = if(exists("data_stress")) data_stress else NULL
  )
  
  # Remove NULL entries
  data_objects <- data_objects[!sapply(data_objects, is.null)]
  
  if (length(data_objects) == 0) {
    cat("No data objects found. Please ensure data is loaded with names like:\n")
    cat("data_cgm, data_hr, data_kcal, data_oxygen, data_rr, data_sleep, data_stress, data_steps\n")
    return(NULL)
  }
  
  # Analyze each modality
  raw_results <- list()
  
  for (modality in names(data_objects)) {
    raw_results[[modality]] <- analyze_data_coverage(data_objects[[modality]], modality, col_info)
  }
  
  # Filter subjects and track removals
  filtering_output <- filter_subjects_by_days(raw_results, MIN_DAYS_THRESHOLD)
  
  return(list(
    raw_results = raw_results,
    filtered_results = filtering_output$filtered_results,
    removal_tracking = filtering_output$removal_tracking
  ))
}

# Function to create summary tables with participant IDs
create_summary_tables <- function(results) {
  
  # Summary table 1: Subjects per modality (before and after filtering)
  subjects_summary <- data.frame(
    Modality = character(),
    Original_N_Subjects = numeric(),
    Filtered_N_Subjects = numeric(),
    Removed_N_Subjects = numeric(),
    Removal_Rate = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Summary table 2: Days per subject per modality (filtered data only)
  days_summary <- data.frame(
    Modality = character(),
    Participant_ID = character(),
    N_Records = numeric(),
    N_Days = numeric(),
    Missing_Percentage = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (modality in names(results$filtered_results)) {
    result <- results$filtered_results[[modality]]
    removal_info <- results$removal_tracking[[modality]]
    
    # Add to subjects summary
    subjects_summary <- rbind(subjects_summary, data.frame(
      Modality = toupper(modality),
      Original_N_Subjects = removal_info$original_count,
      Filtered_N_Subjects = removal_info$filtered_count,
      Removed_N_Subjects = removal_info$removed_count,
      Removal_Rate = removal_info$removal_rate * 100
    ))
    
    # Add to days summary (filtered data only)
    for (subject_result in result$subject_results) {
      days_summary <- rbind(days_summary, data.frame(
        Modality = toupper(modality),
        Participant_ID = subject_result$participant_id,
        N_Records = subject_result$n_records,
        N_Days = ifelse(is.na(subject_result$n_days), 0, subject_result$n_days),
        Missing_Percentage = round(subject_result$missingness$missing_percentage, 2)
      ))
    }
  }
  
  return(list(
    subjects_summary = subjects_summary,
    days_summary = days_summary
  ))
}

# Function to create missingness summary
create_missingness_summary <- function(results) {
  
  missingness_summary <- data.frame(
    Modality = character(),
    Participant_ID = character(),
    Total_Values = numeric(),
    Missing_Values = numeric(),
    Missing_Percentage = numeric(),
    Columns_With_Missing = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Detailed column missingness
  detailed_missingness <- data.frame(
    Modality = character(),
    Participant_ID = character(),
    Column = character(),
    Missing_Count = numeric(),
    Total_Rows = numeric(),
    Missing_Percentage = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (modality in names(results$filtered_results)) {
    result <- results$filtered_results[[modality]]
    
    for (subject_result in result$subject_results) {
      participant_id <- subject_result$participant_id
      missingness <- subject_result$missingness
      
      # Add to summary
      missingness_summary <- rbind(missingness_summary, data.frame(
        Modality = toupper(modality),
        Participant_ID = participant_id,
        Total_Values = missingness$total_values,
        Missing_Values = missingness$missing_values,
        Missing_Percentage = round(missingness$missing_percentage, 2),
        Columns_With_Missing = length(missingness$columns_with_missing)
      ))
      
      # Add detailed column missingness
      if (nrow(missingness$column_missingness) > 0) {
        col_miss_data <- missingness$column_missingness %>%
          mutate(
            Modality = toupper(modality),
            Participant_ID = participant_id
          ) %>%
          select(Modality, Participant_ID, Column = column, 
                 Missing_Count = missing_count, Total_Rows = total_rows,
                 Missing_Percentage = missing_percentage)
        
        detailed_missingness <- rbind(detailed_missingness, col_miss_data)
      }
    }
  }
  
  return(list(
    summary = missingness_summary,
    detailed = detailed_missingness
  ))
}

# Function to create modality-level missingness statistics
create_modality_missingness_stats <- function(missingness_data) {
  
  modality_stats <- missingness_data$summary %>%
    group_by(Modality) %>%
    summarise(
      N_Participants = n(),
      Mean_Missing_Percentage = round(mean(Missing_Percentage, na.rm = TRUE), 2),
      Median_Missing_Percentage = round(median(Missing_Percentage, na.rm = TRUE), 2),
      Min_Missing_Percentage = round(min(Missing_Percentage, na.rm = TRUE), 2),
      Max_Missing_Percentage = round(max(Missing_Percentage, na.rm = TRUE), 2),
      SD_Missing_Percentage = round(sd(Missing_Percentage, na.rm = TRUE), 2),
      Total_Missing_Values = sum(Missing_Values, na.rm = TRUE),
      Total_Values = sum(Total_Values, na.rm = TRUE),
      Overall_Missing_Percentage = round((sum(Missing_Values, na.rm = TRUE) / sum(Total_Values, na.rm = TRUE)) * 100, 2),
      .groups = 'drop'
    ) %>%
    arrange(desc(Overall_Missing_Percentage))
  
  return(modality_stats)
}
```

```{r run-analysis}
# Run the analysis
results <- run_coverage_analysis()

if (!is.null(results)) {
  summary_tables <- create_summary_tables(results)
  subjects_df <- summary_tables$subjects_summary
  days_df <- summary_tables$days_summary
  
  # Create missingness analysis
  missingness_data <- create_missingness_summary(results)
  modality_missingness_stats <- create_modality_missingness_stats(missingness_data)
  
  # Create date range analysis
  date_range_summary <- create_date_range_summary(results)
  temporal_overlap <- create_temporal_overlap_analysis(date_range_summary)
  
  # Calculate summary statistics (filtered data)
  days_summary_stats <- days_df %>%
    filter(N_Days > 0) %>%
    group_by(Modality) %>%
    summarise(
      N_Subjects_with_Data = n(),
      Mean_Days = mean(N_Days, na.rm = TRUE),
      Median_Days = median(N_Days, na.rm = TRUE),
      Min_Days = min(N_Days, na.rm = TRUE),
      Max_Days = max(N_Days, na.rm = TRUE),
      SD_Days = sd(N_Days, na.rm = TRUE),
      Mean_Missing_Percentage = mean(Missing_Percentage, na.rm = TRUE),
      Median_Missing_Percentage = median(Missing_Percentage, na.rm = TRUE),
      .groups = 'drop'
    )
}
```

## Data Summary

::: {.panel-tabset}

### Filtering Summary

#### Subject Removal Summary
```{r filtering-table}
if (!is.null(results)) {
  subjects_df %>%
    arrange(desc(Removal_Rate)) %>%
    mutate(Removal_Rate = paste0(round(Removal_Rate, 1), "%")) %>%
    kable(col.names = c("Modality", "Original Subjects", "Filtered Subjects", 
                        "Removed Subjects", "Removal Rate"),
          caption = paste0("Subject filtering summary (minimum ", MIN_DAYS_THRESHOLD, " days requirement)")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
    column_spec(5, color = ifelse(as.numeric(gsub("%", "", subjects_df$Removal_Rate)) > 50, "red", "black"))
}
```

### Date Range Analysis

#### Date Coverage Summary by Participant
```{r date-range-table}
if (!is.null(results) && exists("date_range_summary")) {
  date_range_summary %>%
    arrange(Modality, Participant_ID) %>%
    mutate(
      Start_Date = as.character(Start_Date),
      End_Date = as.character(End_Date),
      Coverage_Percentage = ifelse(is.na(Coverage_Percentage), "N/A", paste0(round(Coverage_Percentage, 1), "%"))
    ) %>%
    datatable(
      caption = "Date ranges and coverage for each participant by modality (filtered data)",
      options = list(
        pageLength = 15,
        scrollX = TRUE,
        searchHighlight = TRUE
      ),
      filter = 'top'
    ) %>%
    formatStyle(
      'N_Days',
      background = styleColorBar(range(date_range_summary$N_Days, na.rm = TRUE), 'lightblue'),
      backgroundSize = '98% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'center'
    ) %>%
    formatStyle(
      'Date_Gaps',
      color = styleInterval(c(0, 5, 15), c('green', 'orange', 'red', 'darkred'))
    )
}
```

#### Temporal Coverage Statistics
```{r temporal-stats-table}
if (!is.null(results) && exists("date_range_summary")) {
  temporal_stats <- date_range_summary %>%
    filter(!is.na(Start_Date)) %>%
    group_by(Modality) %>%
    summarise(
      N_Participants = n(),
      Earliest_Start = min(Start_Date, na.rm = TRUE),
      Latest_End = max(End_Date, na.rm = TRUE),
      Mean_Duration = round(mean(Duration_Days, na.rm = TRUE), 1),
      Median_Duration = round(median(Duration_Days, na.rm = TRUE), 1),
      Mean_Coverage = round(mean(Coverage_Percentage, na.rm = TRUE), 1),
      Mean_Gaps = round(mean(Date_Gaps, na.rm = TRUE), 1),
      .groups = 'drop'
    ) %>%
    mutate(
      Study_Span_Days = as.numeric(Latest_End - Earliest_Start) + 1,
      Earliest_Start = as.character(Earliest_Start),
      Latest_End = as.character(Latest_End)
    )
  
  temporal_stats %>%
    kable(col.names = c("Modality", "N Participants", "Earliest Start", "Latest End", 
                        "Mean Duration (Days)", "Median Duration (Days)", 
                        "Mean Coverage (%)", "Mean Gaps", "Study Span (Days)"),
          caption = "Temporal coverage statistics by modality") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
}
```

### Missingness Analysis

#### Modality-Level Missingness Statistics
```{r modality-missingness-table}
if (!is.null(results)) {
  modality_missingness_stats %>%
    kable(col.names = c("Modality", "N Participants", "Mean Missing %", "Median Missing %", 
                        "Min Missing %", "Max Missing %", "SD Missing %", 
                        "Total Missing Values", "Total Values", "Overall Missing %"),
          caption = "Missingness statistics by modality (filtered data)") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
    column_spec(10, color = ifelse(modality_missingness_stats$Overall_Missing_Percentage > 20, "red", 
                                  ifelse(modality_missingness_stats$Overall_Missing_Percentage > 10, "orange", "green")))
}
```

#### Participant-Level Missingness Summary
```{r participant-missingness-table}
if (!is.null(results)) {
  missingness_data$summary %>%
    arrange(Modality, desc(Missing_Percentage)) %>%
    datatable(
      caption = "Missingness by participant and modality (filtered data)",
      options = list(
        pageLength = 15,
        scrollX = TRUE,
        searchHighlight = TRUE
      ),
      filter = 'top'
    ) %>%
    formatStyle(
      'Missing_Percentage',
      background = styleColorBar(range(missingness_data$summary$Missing_Percentage), 'lightcoral'),
      backgroundSize = '98% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'center'
    ) %>%
    formatStyle(
      'Missing_Percentage',
      color = styleInterval(c(10, 20, 50), c('black', 'darkorange', 'red', 'darkred'))
    )
}
```

### Summary Tables

#### Subjects per Modality (After Filtering)
```{r subjects-table}
if (!is.null(results)) {
  subjects_df %>%
    select(Modality, Filtered_N_Subjects) %>%
    arrange(desc(Filtered_N_Subjects)) %>%
    kable(col.names = c("Data Modality", "Number of Subjects (≥5 days)"),
          caption = "Number of subjects with sufficient data in each modality") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
}
```

#### Days Summary Statistics (Filtered Data)
```{r days-summary-table}
if (!is.null(results)) {
  days_summary_stats %>%
    mutate(across(where(is.numeric), ~round(.x, 1))) %>%
    kable(col.names = c("Modality", "Subjects with Data", "Mean Days", "Median Days", 
                        "Min Days", "Max Days", "SD Days", "Mean Missing %", "Median Missing %"),
          caption = "Summary statistics for days of data per subject by modality (filtered data)") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
}
```

### Detailed Data

#### Interactive Subject-Level Data (Filtered)
```{r detailed-table}
if (!is.null(results)) {
  days_df %>%
    filter(N_Days > 0) %>%
    arrange(Modality, desc(N_Days)) %>%
    datatable(
      caption = "Detailed view: Days
      of data per subject by modality (filtered data)",
      options = list(
        pageLength = 15,
        scrollX = TRUE,
        searchHighlight = TRUE,
        dom = 'Bfrtip',
        buttons = c('copy', 'csv', 'excel')
      ),
      filter = 'top'
    ) %>%
    formatStyle(
      'N_Days',
      background = styleColorBar(range(days_df$N_Days[days_df$N_Days > 0]), 'lightblue'),
      backgroundSize = '98% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'center'
    ) %>%
    formatStyle(
      'Missing_Percentage',
      background = styleColorBar(range(days_df$Missing_Percentage), 'lightcoral'),
      backgroundSize = '98% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'center'
    )
}
```

:::

## Visualizations

::: {.panel-tabset}

### Subject Distribution

```{r subject-distribution-plot}
if (!is.null(results)) {
  p1 <- subjects_df %>%
    arrange(desc(Filtered_N_Subjects)) %>%
    ggplot(aes(x = reorder(Modality, Filtered_N_Subjects), y = Filtered_N_Subjects)) +
    geom_col(fill = "steelblue", alpha = 0.8) +
    geom_text(aes(label = Filtered_N_Subjects), hjust = -0.1, size = 3.5) +
    coord_flip() +
    labs(
      title = paste0("Number of Subjects per Modality (≥", MIN_DAYS_THRESHOLD, " days)"),
      x = "Data Modality",
      y = "Number of Subjects"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 12)
    )
  
  print(p1)
}
```

### Days Distribution

```{r days-distribution-plot}
if (!is.null(results)) {
  # Box plot of days distribution
  p2 <- days_df %>%
    filter(N_Days > 0) %>%
    ggplot(aes(x = reorder(Modality, N_Days, median), y = N_Days, fill = Modality)) +
    geom_boxplot(alpha = 0.7, outlier.alpha = 0.5) +
    geom_jitter(width = 0.2, alpha = 0.6, size = 1) +
    coord_flip() +
    labs(
      title = "Distribution of Days of Data per Subject by Modality",
      x = "Data Modality",
      y = "Number of Days"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      axis.text = element_text(size = 11),
      axis.title = element_text(size = 12),
      legend.position = "none"
    ) +
    scale_fill_viridis_d()
  
  print(p2)
}
```

:::

## Column-Level Analysis

::: {.panel-tabset}

### Column-Level Missingness

```{r column-missingness-table}
if (!is.null(results)) {
  # Show detailed column missingness for columns with >0% missing
  high_missingness_cols <- missingness_data$detailed %>%
    filter(Missing_Percentage > 0) %>%
    arrange(desc(Missing_Percentage))
  
  if (nrow(high_missingness_cols) > 0) {
    high_missingness_cols %>%
      datatable(
        caption = "Column-level missingness analysis (columns with missing data only)",
        options = list(
          pageLength = 15,
          scrollX = TRUE,
          searchHighlight = TRUE
        ),
        filter = 'top'
      ) %>%
      formatStyle(
        'Missing_Percentage',
        background = styleColorBar(range(high_missingness_cols$Missing_Percentage), 'lightcoral'),
        backgroundSize = '98% 88%',
        backgroundRepeat = 'no-repeat',
        backgroundPosition = 'center'
      ) %>%
      formatStyle(
        'Missing_Percentage',
        color = styleInterval(c(25, 50, 75), c('black', 'darkorange', 'red', 'darkred'))
      )
  } else {
    cat("No columns with missing data found in the filtered dataset.")
  }
}
```

:::

## Cross-Modality Analysis

::: {.panel-tabset}

### Participant Overlap Matrix

```{r overlap-matrix}
if (!is.null(results)) {
  # Create participant overlap matrix
  participant_modality_matrix <- days_df %>%
    filter(N_Days > 0) %>%
    select(Modality, Participant_ID) %>%
    mutate(Has_Data = 1) %>%
    pivot_wider(names_from = Modality, values_from = Has_Data, values_fill = 0)
  
  if (nrow(participant_modality_matrix) > 0) {
    participant_modality_matrix %>%
      datatable(
        caption = "Participant data availability matrix (1 = has data, 0 = no data)",
        options = list(
          pageLength = 15,
          scrollX = TRUE,
          searchHighlight = TRUE
        ),
        filter = 'top'
      ) %>%
      formatStyle(
        columns = names(participant_modality_matrix)[-1],
        backgroundColor = styleEqual(c(0, 1), c('lightcoral', 'lightgreen'))
      )
  }
}
```

### Multi-Modality Coverage Summary

```{r multi-modality-summary}
if (!is.null(results)) {
  # Calculate how many modalities each participant has data for
  modality_coverage <- days_df %>%
    filter(N_Days > 0) %>%
    group_by(Participant_ID) %>%
    summarise(
      N_Modalities = n(),
      Modalities_List = paste(Modality, collapse = ", "),
      Total_Days = sum(N_Days),
      Mean_Days_Per_Modality = round(mean(N_Days), 1),
      .groups = 'drop'
    ) %>%
    arrange(desc(N_Modalities), desc(Total_Days))
  
  modality_coverage %>%
    datatable(
      caption = "Multi-modality data coverage per participant",
      options = list(
        pageLength = 15,
        scrollX = TRUE,
        searchHighlight = TRUE
      ),
      filter = 'top'
    ) %>%
    formatStyle(
      'N_Modalities',
      background = styleColorBar(range(modality_coverage$N_Modalities), 'lightblue'),
      backgroundSize = '98% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'center'
    )
}
```

:::

## Summary

```{r final-summary}
if (!is.null(results)) {
  cat("## Key Findings\n\n")
  
  # Modality with best coverage
  best_modality <- subjects_df$Modality[which.max(subjects_df$Filtered_N_Subjects)]
  best_count <- max(subjects_df$Filtered_N_Subjects)
  
  cat(sprintf("- **Modality with highest subject count:** %s (%d subjects)\n", best_modality, best_count))
  
  # Modality with worst coverage
  worst_modality <- subjects_df$Modality[which.min(subjects_df$Filtered_N_Subjects)]
  worst_count <- min(subjects_df$Filtered_N_Subjects)
  
  cat(sprintf("- **Modality with lowest subject count:** %s (%d subjects)\n\n", worst_modality, worst_count))
  
  # Missingness insights
  if (exists("modality_missingness_stats")) {
    highest_missing <- modality_missingness_stats$Modality[which.max(modality_missingness_stats$Overall_Missing_Percentage)]
    highest_missing_pct <- max(modality_missingness_stats$Overall_Missing_Percentage)
    
    lowest_missing <- modality_missingness_stats$Modality[which.min(modality_missingness_stats$Overall_Missing_Percentage)]
    lowest_missing_pct <- min(modality_missingness_stats$Overall_Missing_Percentage)
    
    cat(sprintf("- **Highest data missingness:** %s (%.1f%% missing)\n", highest_missing, highest_missing_pct))
    cat(sprintf("- **Lowest data missingness:** %s (%.1f%% missing)\n\n", lowest_missing, lowest_missing_pct))
  }
}
```

*Analysis generated on `r Sys.Date()`*